Bayesian latent class model with sparse correlation for etiology estimation

### 2017-01-30 Reports/draft.v1/draft_latentquadexp.tex
figure caption in draft

\begin{table}[h]
\centering
\caption{The common hyper-parameters used for model fitting in simulation studies}
\label{tab:study3}
\begin{tabular}{lrrrrrrrr}
\hline
          & $a_k$    & $b_k$     & $c_k$     & $d_k$     & $e_k$ & $f_k$ & $\sigma_1$  & $\sigma_2$    \\ \hline
Study I   & 7.6 & 59  & 5  & 3  & 1 & 1 & 0.2 & 0.2 \\
Study II  & 50 & 10  & 12  & 1  & 1 & 12 & 0.2 & 0.2 \\
Study III & 7.6 & 59  & 5  & 3  & 1 & 1 & 0.2 & 0.2 \\ \hline
\end{tabular}
\end{table}

@@

\begin{center}
\includegraphics[scale=0.6]{StudyI.png}
\captionof{figure}{{\bf Overall estimation accuracy of models in study I}: }
\label{fig:study1}
\end{center}

Model performances in Study II is summarized in table \ref{tab: } and plotted in figure \ref{}. The figure shows that when the true data generating mechanism is actually single pathogen infection only, the pLCM-1 has the best estimation accuracy since that is the true model, and the pLCM-2 performs the worst because it tends to attribute the cause of disease to two-pathogen infection. The LSC model, on the other hand, shows an increasing trend in estimation accuracy as the prior mean of $\theta_2$ gets more negative, because $\theta_2 = -\infty$ makes the model equivalent to the true model. And in practice, as we can see in figure \ref{}, if the true model is not known, the LSC model can still provide an estimate almost as accurate as the pLCM-1 with a moderately negative (e.g. between $-10$ and $-5$) prior mean of $\theta_2$.\\

It is very intuitive to think that better measurement quality leads to more accurate etiology estimation. Figure \ref{} , which compares the etiology estimations, $\hat{\mathbb{E}}(L|Y=1,X)$, from study III and two scenarios of study I, confirms this hypothesis. In this figure, the sampling distributions of the etiology estimations for pathogen A, B and C in the upper panel have larger standard deviations than those in the mid panel, their shapes are more skewed and the estimation bias tends to be larger, and such distributions for pathogen D and E are almost the same in the upper and mid panel. Comparing the lower panel to the mid panel, it is clear that the sampling distributions estimated from high quality measurements have much smaller standard deviations and basically no estimation bias.\\

\begin{center}
\includegraphics[scale=0.45]{DataDescription.png}
\captionof{figure}{{\bf Data Description}: }
\label{}
\end{center}

\section{PERCH Data Analysis}
As what has been introduced in the first section of this paper, the PERCH study enrolled about 4200 children hospitalized for severe/very severe pneumonia and approximately 5300 controls randomly select from communities across $7$ sites around the world. To demonstrate the application of the LSC model for the analysis of PERCH study data, only the Kenya site data, where there is good availability of both BS and SS measurement data, is used so that the site-specific effect if not a concern.

>>>>

\begin{table}[h]
\centering
\caption{The common hyper-parameters used for model fitting in simulation studies}
\label{tab:study3}
\begin{tabular}{lrrrrrrrr}
\hline
          & $a_k$    & $b_k$     & $c_k$     & $d_k$     & $e_k$ & $f_k$ & $\sigma_1$  & $\sigma_2$    \\ \hline
Study I   & 7.6 & 59  & 5  & 3  & 1 & 1 & 2.2 & 2.2 \\
Study II  & 50 & 10  & 12  & 1  & 1 & 1 & 2.2 & 2.2 \\
Study III & 7.6 & 59  & 5  & 3  & 1 & 1 & 2.2 & 2.2 \\ \hline
\end{tabular}
\end{table}

@@

\begin{center}
\includegraphics[scale=0.6]{StudyI.png}
\captionof{figure}{{\bf Overall estimation accuracy of models in study I}: The x axis stands for $\alpha_2$, the prior mean of $\theta_2$, and the y axis represents the Bhattacharyya coefficient (BC). The round dots are values generated by LSC models with different priors. The color of dots indicates the mean of the prior distribution of $p$ and the size of dots indicate the standard deviation of the prior distribution of $p$. These two values are calculated based on $\lambda_a$ and $\lambda_b$. The two horizontal lines are benchmark values generated by pLCM models. The solid line represents the pLCM-1 model which only allows singleton infections, and the dash line represents the pLCM-2 model which allows singleton and all the pair infections.}
\label{fig:study1}
\end{center}

Model performances in Study II is summarized in table \ref{tab:result2} and plotted in figure \ref{fig:study2}. The figure shows that when the true data generating mechanism is actually single pathogen infection only, the pLCM-1 has the best estimation accuracy since that is the true model, and the pLCM-2 performs the worst because it tends to attribute the cause of disease to two-pathogen infection. The LSC model, on the other hand, shows an increasing trend in estimation accuracy as the prior mean of $\theta_2$ gets more negative, because $\theta_2 = -\infty$ makes the model equivalent to the true model. And in practice, as we can see in figure \ref{fig:study2}, if the true model is not known, the LSC model can still provide an estimate almost as accurate as the pLCM-1 with a moderately negative (e.g. between $-10$ and $-5$) prior mean of $\theta_2$.\\

\begin{table}[]
\centering
\caption{Summary of the overall parameter estimation accuracy of each model fitted in study II}
\label{tab:result2}
\begin{tabular}{llll}
\hline
$\alpha_2$ & $\lambda_a$ & $\lambda_b$ & Bhattacharyya \\ \hline
-16        & 6           & 2           & 0.9925        \\
-10        & 6           & 2           & 0.9828        \\
-16        & 4           & 4           & 0.9761        \\
-7.5       & 15          & 5           & 0.9742        \\
-10        & 15          & 5           & 0.974         \\
-7.5       & 6           & 2           & 0.9725        \\
-10        & 4           & 4           & 0.9717        \\
-16        & 15          & 5           & 0.9635        \\
-7.5       & 4           & 4           & 0.9603        \\
-4         & 6           & 2           & 0.9501        \\
-4         & 15          & 5           & 0.9257        \\
-1         & 15          & 5           & 0.9101        \\
-1         & 6           & 2           & 0.9089        \\
-1         & 4           & 4           & 0.9086        \\
-4         & 4           & 4           & 0.9052        \\ \hline
\multicolumn{3}{l}{pLCM-1}             & 0.9981        \\
\multicolumn{3}{l}{pLCM-2}             & 0.8203        \\ \hline
\end{tabular}
\end{table}

\begin{center}
\includegraphics[scale=0.6]{StudyII.png}
\captionof{figure}{{\bf Overall estimation accuracy of models in study II}: The x axis stands for $\alpha_2$, the prior mean of $\theta_2$, and the y axis represents the Bhattacharyya coefficient (BC). The round dots are values generated by LSC models with different priors. The color of dots indicates the mean of the prior distribution of $p$ and the size of dots indicate the standard deviation of the prior distribution of $p$. These two values are calculated based on $\lambda_a$ and $\lambda_b$. The two horizontal lines are benchmark values generated by pLCM models. The solid line represents the pLCM-1 model which only allows singleton infections, and the dash line represents the pLCM-2 model which allows singleton and all the pair infections.}
\label{fig:study2}
\end{center}

It is very intuitive to think that better measurement quality leads to more accurate etiology estimation. Figure \ref{fig:dataquality} , which compares the etiology estimations, $\hat{\mathbb{E}}(L|Y=1,X)$, from study III and two scenarios of study I, confirms this hypothesis. In this figure, the sampling distributions of the etiology estimations for pathogen A, B and C in the upper panel have larger standard deviations than those in the mid panel, their shapes are more skewed and the estimation bias tends to be larger, and such distributions for pathogen D and E are almost the same in the upper and mid panel. Comparing the lower panel to the mid panel, it is clear that the sampling distributions estimated from high quality measurements have much smaller standard deviations and basically no estimation bias.\\

\begin{center}
\includegraphics[scale=0.75]{DataQualityImpact.png}
\captionof{figure}{{\bf The impact of data quality on the LSC model}: The results from study I are plotted in the mid panel, the results from study I data with SS measurements of pathogen A, B, and C removed are plotted in the upper panel, and the results from study III are plotted in the lower panel. In each plot, there are five facets, with each one corresponds to a pathogen. In each facet, the x axis represents four strata that are determined by the two binary covariates, and the y axis stands for the estimated value of etiology probability, $\hat{\mathbb{E}}(L|Y = 1, X)$. The violin shape indicates the estimated density function of the sampling distributions of $\hat{\mathbb{E}}(L|Y = 1, X)$. The three horizontal lines in each violin shape represent the $2.5th$, $50th$, and $97.5th$ percentiles respectively. The red dots show the true values of the corresponding parameters.}
\label{fig:dataquality}
\end{center}

\newpage
\section{PERCH Data Analysis}
As what has been introduced in the first section of this paper, the PERCH study enrolled about 4200 children hospitalized for severe/very severe pneumonia and approximately 5300 controls randomly select from communities across $7$ sites around the world. To demonstrate the application of the LSC model for the analysis of PERCH study data, only the Kenya site data, where there is good availability of both BS and SS measurement data, is used so that the site-specific effect if not a concern. We picked the top $5$ pathogens reported in \cite{wu2015partially} as our candidate pathogens in this analysis. These pathogens are streptococcus pneumoniae (PNEU), haemophilus influenzae (HINF), human metapneumovirus type A or B (HMPV\_A\_B), rhonovirus (RHINO), and respiratory syncytial virus type A or B (RSV). The BS measurements (nasopharyngeal specimen with PCR detection of pathogens - NPPCR) are available for all $281$ cases and $1138$ frequency-matched controls on all $5$ pathogens. The SS measurements (blood culture results - BCX) are only available for all cases on the two bacteria pathogens: PNEU and HINF.\\

Prior scientific knowledge \cite{murdoch2012laboratory} suggests that the TPR of BS measurements (NPPCR) is in the range of $50$\% - $99$\%, and the TPR of SS measurements (BCX) is in the range of $5$\% - $20$\%. Thus we set the hyper-parameters $a_k = 7.6$, $b_k = 59.0$, $c_k = 6.0$, and $d_k = 1.3$ by range matching, and $e_k = f_k = 1$ for non-informativeness. Other hyper-parameter values used for PERCH data analysis are listed in table \ref{tab:perchhyper}.

\begin{table}[h]
\centering
\caption{The hyper-parameters used for the LSC model in PERCH data analysis}
\label{tab:perchhyper}
\begin{tabular}{lrrrrr}
\hline
        & $\alpha_2$ & $\lambda_a$ & $\lambda_b$ & $\sigma_1$  & $\sigma_2$    \\ \hline
LSC     & -5 & 6  & 2 & 2.2 & 2.2 \\ \hline
\end{tabular}
\end{table}
